---
title: "Long_requets"
format: html
editor: visual
---

```{r load libraries}
#| output: false
#| cache: false

library(openalexR)
library(dplyr)
library(jsonlite)
library(tools)
library(stringr)
library(purrr)
library(tidyr)
library(broom)
library(tidyverse)
library(httr)
library(jsonlite)
```

# Introduction

This document contains all the chunks of code which are very long to compute because they make many requests to the API servers of OpenAlex or OpenCitations or necessitate the processing of the big files created by these requests. The requests have been run once, and the result saved in .rds files. These .rds files are read in the main quarto document.

I copy the chunks here in the same layout as the main document.

# Replicating the bibliometric study with data from OpenAlex

## Data extraction from OpenAlex

### Articles' data extraction

#### Using the openalexR package

The following chunk downloads **all the articles' metadata** listed in OpenAlex for the 'Archaeology' subfield which have been published between 1975 and 2025. This takes a very long time and creates big files.

```{r Downloading data for 'Archaeology' articles in OpenAlex (AQ)}
#| eval: false

# This data extraction takes several hours.

# Here we extract the articles in the subfield Archaeology 
works_archaeology_search_1975_1985 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204", # 1204 is the id of the subfield Archaeology
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "1975-01-01",
  to_publication_date = "1985-12-31",
  verbose = TRUE)

works_archaeology_search_1986_1995 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "1986-01-01",
  to_publication_date = "1995-12-31",
  verbose = TRUE
)

works_archaeology_search_1996_2000 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "1996-01-01",
  to_publication_date = "2000-12-31",
  verbose = TRUE
)

works_archaeology_search_2001_2003 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2001-01-01",
  to_publication_date = "2003-12-31",
  verbose = TRUE
)

works_archaeology_search_2004_2007 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2004-01-01",
  to_publication_date = "2007-12-31",
  verbose = TRUE
)

works_archaeology_search_2008_2011 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2008-01-01",
  to_publication_date = "2011-12-31",
  verbose = TRUE
)

works_archaeology_search_2012_2015 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2012-01-01",
  to_publication_date = "2015-12-31",
  verbose = TRUE
)

works_archaeology_search_2016_2019 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2016-01-01",
  to_publication_date = "2019-12-31",
  verbose = TRUE
)

works_archaeology_search_2020_2022 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2020-01-01",
  to_publication_date = "2022-12-31",
  verbose = TRUE
)

works_archaeology_search_2023_2024 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2023-01-01",
  to_publication_date = "2024-12-31",
  verbose = TRUE
)

works_archaeology_search_2025 <- oa_fetch(
  entity = "works",
  type = "article",
  topics.subfield.id = "1204",
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2025-01-01",
  to_publication_date = "2025-12-31",
  verbose = TRUE
)

# Combining the data together
works_archaeology_1975_2025 <- rbind(works_archaeology_search_1975_1985,works_archaeology_search_1986_1995,works_archaeology_search_1996_2000,works_archaeology_search_2001_2003,works_archaeology_search_2004_2007,works_archaeology_search_2008_2011,works_archaeology_search_2012_2015,works_archaeology_search_2016_2019,works_archaeology_search_2020_2022,works_archaeology_search_2023_2024,works_archaeology_search_2025)

saveRDS(works_archaeology_1975_2025, "data/OpenAlex_works_archaeology_1975_2025.rds")

# Define the variables we want to keep in the shorter version of this 1.8 million-lines dataframe
variables_to_keep <- c("id", "title", "authorships","doi","publication_year","source_display_name","host_organization_name","referenced_works_count","first_page","last_page","abstract","referenced_works")

# Apply the selection of variables to the list of papers extracted from OpenAlex
# and cleaning it (titles, page numbers, preparing authors' lists, removing duplicates)

works_archaeology_short_1975_2025 <- works_archaeology_1975_2025 %>% select(all_of(variables_to_keep)) %>% 
  mutate(title = str_trim(title)) %>% # delete double space
  mutate(title = str_to_lower(title)) %>% # put titles to lower case
  mutate(title = str_remove(title, "<i>")) %>%
  mutate(title = str_remove(title, "</i>")) %>% 
  mutate(referenced_works = str_remove(referenced_works, "\n")) %>%
  mutate(authors = map_chr(authorships, ~ paste(.x$display_name, collapse = ", "))) %>%
  distinct(title, .keep_all = TRUE)  # removes works present more than once based on title

saveRDS(works_archaeology_short_1975_2025, "data/OpenAlex_works_archaeology_short_1975_2025.rds")

# Keeping only the articles from the OpenAlex top25 2-year mean citedness journals
works_archaeology_top25_1975_2025 <- works_archaeology_short_1975_2025 %>% filter(source_display_name %in% jci_top_25_2ymc$display_name)

# Grouping by source to know how many journals in archaeology
sources <- works_archaeology_short_1975_2025 %>%
  group_by(source_display_name) %>%
  summarise(n = n(), .groups = "drop")

sources_100articles <- sources %>% filter(n>99)
n_sources_100articles <- nrow(sources_100articles)
sources_500articles <- sources %>% filter(n>499)
n_sources_500articles <- nrow(sources_500articles)
sources_1000articles <- sources %>% filter(n>999)
n_sources_1000articles <- nrow(sources_1000articles)
```

The following chunk downloads the metadata **only for the articles published in the top-25-2ymc**. This takes only few minutes.

```{r Downloading OpenAlex works from top25 2ymc journals (AQ)}
#| eval: false

# Time consuming (few minutes) but result saved in .rds

# Below the code download several batch of data only for the top25 journals identified by their 2ymc value so that it is too big, and then bind them all together in a single tibble with the necessary variables and not all of them.

works_search_1975_1985 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "1975-01-01",
  to_publication_date = "1985-12-31",
  verbose = TRUE)

works_search_1986_1995 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "1986-01-01",
  to_publication_date = "1995-12-31",
  verbose = TRUE
)

works_search_1996_2000 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "1996-01-01",
  to_publication_date = "2000-12-31",
  verbose = TRUE
)

works_search_2001_2003 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2001-01-01",
  to_publication_date = "2003-12-31",
  verbose = TRUE
)

works_search_2004_2007 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2004-01-01",
  to_publication_date = "2007-12-31",
  verbose = TRUE
)

works_search_2008_2011 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2008-01-01",
  to_publication_date = "2011-12-31",
  verbose = TRUE
)

works_search_2012_2015 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2012-01-01",
  to_publication_date = "2015-12-31",
  verbose = TRUE
)

works_search_2016_2019 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2016-01-01",
  to_publication_date = "2019-12-31",
  verbose = TRUE
)

works_search_2020_2023 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2020-01-01",
  to_publication_date = "2023-12-31",
  verbose = TRUE
)

works_search_2024_2025 <- oa_fetch(
  entity = "works",
  type = "article",
  journal = id_25_2ymc, # id is the list of top25 2ymc journals OpenAlex ids
  count_only = FALSE,
  output = "dataframe",
  from_publication_date = "2024-01-01",
  to_publication_date = "2025-12-31",
  verbose = TRUE
)

# Combining the data together
liste_tibbles <- list(works_search_1975_1985,works_search_1986_1995,works_search_1996_2000,works_search_2001_2003,works_search_2004_2007,works_search_2008_2011,works_search_2012_2015,works_search_2016_2019,works_search_2020_2023,works_search_2024_2025)


# Define the variables we want to keep
variables_to_keep <- c("id", "title", "authorships","doi","publication_year","source_display_name","host_organization_name","referenced_works_count","first_page","last_page","abstract","referenced_works")

# Apply the selection to the list of papers extracted from OpenAlex
# and cleaning it (titles, page numbers, preparing authors' lists, removing duplicates)

Works_top25_1975_2025 <- bind_rows(lapply(liste_tibbles, function(x) {
  x %>% select(all_of(variables_to_keep)) })) %>% 
  mutate(title = str_trim(title)) %>% # delete double space
  mutate(title = str_to_lower(title)) %>% # put titles to lower case
  mutate(title = str_remove(title, "<i>")) %>%
  mutate(title = str_remove(title, "</i>")) %>% 
  mutate(referenced_works = str_remove(referenced_works, "\n")) %>%
  mutate(authors = map_chr(authorships, ~ paste(.x$display_name, collapse = ", "))) %>%
  distinct(title, .keep_all = TRUE)  # removes ca. 1477 works present more than once (based on title)

# Prepare dataframe exactly as in Marwick 2025
items_df = dplyr::tibble(
    authors =         Works_top25_1975_2025$authors,
    authors_n =       map_int(Works_top25_1975_2025$authorships, ~ nrow(.x)),
    title =           Works_top25_1975_2025$title,
    title_n =         sapply(strsplit(Works_top25_1975_2025$title, "\\s+"), length),
    journal =         Works_top25_1975_2025$source_display_name,
    abstract  =       Works_top25_1975_2025$abstract,
    refs =            Works_top25_1975_2025$referenced_works,
    refs_n =          Works_top25_1975_2025$referenced_works_count,
    pages_n =         as.numeric(Works_top25_1975_2025$last_page)-as.numeric(Works_top25_1975_2025$first_page),
    year =            Works_top25_1975_2025$publication_year,
    doi =             Works_top25_1975_2025$doi
  )

items_df = items_df %>% filter(journal %in% jci_top_25_2ymc$display_name)

# save to disk so we can re-use it for the next steps to save time
saveRDS(items_df, "data/OpenAlex_Works_top25_1975_2025.rds")
```

The following chnk download the metadata for all the references cited in every article published between 1975 and 2025 in the top-25-2ymc. this takes more than 20 hours to do.

```{r get year and journals of refs through OpenAlex API (AQ)}
#| label: get year for each ref of each paper
#| eval: false
#very time consuming but result saved in .rds

# Beware this takes ca. 20 hours (45k requests to the API) and should not be run if not really necessary
# Use the result year_refs_list.rds instead
# Thus, I put eval: false to this chunk so that it does not run when rendering the quarto document!

items_df = readRDS("data/OpenAlex_Works_top25_1975_2025.rds")

# Keeping only the OpenAlex ID of each ref in the refs variable
items_df = items_df  %>% 
  mutate(refs = str_remove_all(refs, "https://openalex.org/")) %>%
  mutate(refs = str_remove(refs, "^c\\(")) %>%
  mutate(refs = str_remove_all(refs, '"')) %>%
  mutate(refs = str_remove_all(refs, '\\)')) %>%
  mutate(refs = str_remove_all(refs, '\n'))

# Creating the list before the loop
year_refs_list <- list()
journal_refs_list <- list()

# Loop for getting the list of all refs for each paper and keeping only its publication years 
# (very very long)
for (i in 1:nrow(items_df)) {
year_refs_search <- tryCatch({
  oa_fetch(
  entity = "works",
  identifier = strsplit(items_df$refs[[i]], ", ")[[1]],
  count_only = FALSE,
  output = "dataframe",
  verbose = TRUE,
  mailto = "alain.queffelec@u-bordeaux.fr"
) 
}, error = function(e) {
    message(paste("Error fetching data for row", i, ": ", e$message))
    NULL
  })

if (!is.null(year_refs_search)) {
    year_refs <- year_refs_search$publication_year
    journals <- year_refs_search$source_display_name
    year_refs_list[[i]] <- year_refs
    journal_refs_list[[i]] <- journals
  }
}

# Save the list on the disk
saveRDS(year_refs_list, "data/year_refs_list.rds")
saveRDS(journal_refs_list, "data/journal_refs_list.rds")
```

## Comparison between WoS and OpenAlex

### Articles' comparison

The following chunk of code allows to extract from OpenCitations the number of references cited by each article based on a list of DOIs.

```{r extracting number of references from OpenCitations for articles in OpenAlex and WoS (AQ)}
#| eval: false

# Function to call the API of OpenCitations and extract the number of references for each article identified by its doi
get_references_count <- function(doi) {
  API_CALL <- paste0("https://api.opencitations.net/index/v2/references/doi:", doi,"?format=csv")

  # make the API request
  response <- GET(url = API_CALL)

  # check the status of the response
  if (status_code(response) == 200) {
    data <- content(response, as = "parsed")
    # Extract the number of citations
    references_count <- nrow(data)
    return(references_count)
  } else {
    message(paste("Error for the DOI", doi, ": Code HTTP", status_code(response)))
    return(NA)
  }
}

# List of common dois from OpenAlex and WoS
dois <- Common_Articles_limited_short$doi_OA

# Preparation of batches of request since the number is limited by OpenCitations
batch_size <- 180
total_dois <- length(dois)
total_batches <- ceiling(total_dois / batch_size)

# List to store the results
results <- data.frame()

# Loop to request the references by batches
for (i in 1:total_batches) {
  start_idx <- (i - 1) * batch_size + 1
  end_idx <- min(i * batch_size, total_dois)

  current_batch <- dois[start_idx:end_idx]

  # Request for each doi of the batch
  batch_results <- as.data.frame(sapply(current_batch, get_references_count))

  # Merge this batch with previous batches
  results <- c(results, batch_results)

  # Display of the current step
  cat(paste("Batch", i, "in", total_batches, "just treated.\n"))

  # Pause of 70 s between batches since OpenCitations limit the number of request per minute
  if (i < total_batches) {
    cat("70s pause  before next batch.\n")
    Sys.sleep(70)
  }
}

# Convert results in a dataframe
refs_OpenCitations <- do.call(rbind, lapply(results, as.data.frame))
refs_OpenCitations$doi = Common_Articles_limited_short$doi_OA
colnames(refs_OpenCitations) = c("refs_n_OC","doi")
rownames(refs_OpenCitations) <- NULL

# Save the results on the disk
saveRDS(refs_OpenCitations, "data/refs_OpenCitations.rds")
```

The following chunk of code allows to extract from OpenCitations the metadata of articles based on a list of DOIs.

```{r extracting informations from OpenCitations for articles in both OpenAlex and WoS (AQ)}
#| eval: false
#| cache: TRUE

# Function to call the API of OpenCitations and extract the number of references for each article identified by its doi
get_metadata <- function(doi) {
  API_CALL <- paste0("https://api.opencitations.net/meta/v1/metadata/doi:", doi,"?format=csv")

  # make the API request
  response <- GET(url = API_CALL)

  # check the status of the response
  if (status_code(response) == 200) {
    metadata <- content(response, as = "parsed")
    return(metadata)
  } else {
    message(paste("Error for the DOI", doi, ": Code HTTP", status_code(response)))
    return(NA)
  }
}

# List of common dois from OpenAlex and WoS
dois <- Common_Articles_limited_short$doi_OA

# Preparation of batches of request since the number is limited by OpenCitations
batch_size <- 180
total_dois <- length(dois)
total_batches <- ceiling(total_dois / batch_size)

# List to store the results
results <- list()

# Loop to request the references by batches
for (i in 1:total_batches) {
  start_idx <- (i - 1) * batch_size + 1
  end_idx <- min(i * batch_size, total_dois)

  current_batch <- dois[start_idx:end_idx]

  # Request for each doi of the batch
  batch_results <- as.data.frame(sapply(current_batch, get_metadata))

  # Merge this batch with previous batches
  results <- c(results, batch_results)

  # Display of the current step
  cat(paste("Batch", i, "in", total_batches, "just treated.\n"))

  # Pause of 70 s between batches since OpenCitations limit the number of request per minute
  if (i < total_batches) {
    cat("70s pause  before next batch.\n")
    Sys.sleep(70)
  }
}

#keep only the year for the pub_date
results_year <- lapply(results, function(obj) {
  obj$pub_date <- substr(obj$pub_date, 1, 4)
  return(obj)
})

# Convert results in a dataframe
infos_OpenCitations <- do.call(rbind, lapply(results_year, as.data.frame))
infos_OpenCitations$doi_OC = row.names(infos_OpenCitations)
rownames(infos_OpenCitations) <- NULL

# Prepare data for comparison title_n_OC etc.
infos_OpenCitations$title_n_OC = sapply(strsplit(infos_OpenCitations$title, "\\s+"), length)
infos_OpenCitations$authors_n_OC = sapply(infos_OpenCitations$author, function(x) str_count(x, fixed("[")))
infos_OpenCitations = infos_OpenCitations %>%
  separate(col = page, into = c("first_page", "last_page"), sep = "-") %>%
  mutate(
    first_page = as.numeric(first_page),
    last_page = as.numeric(last_page),
    pages_n_OC = last_page - first_page + 1
  )
infos_OpenCitations$year_OC = as.numeric(infos_OpenCitations$pub_date)

# Save the results on the disk
saveRDS(infos_OpenCitations, "data/infos_OpenCitations.rds")
```